## Exploratory Data Analysis

```{r, include=FALSE}
train <- read.csv("../data/raw/bank_marketing_train.csv")
```

```{r}
datatable(head(train, 100), options = list(scrollX = TRUE))
str(train)
attach(train)
```

### Variable descriptions
#### Bank client data:
1. `X` (Integer): id of customer
2. `age` (Integer): age of the customer
3. `job` (Categorical): occupation
4. `marital` (Categorical): marital status
5. `education` (Categorical): education level
6. `default` (Binary): has credit in default?
7. `housing` (Binary): has housing loan?
8. `loan` (Binary): has personal loan?
9. `contact` (Categorical): contact communication type
10. `month` (Categorical): last contact month of year
11. `day_of_week` (Integer): last contact day of the week
12. `duration` (Integer): last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model

#### Other attributes:
13. `campaign` (Integer): number of contacts performed during this campaign and for this client (numeric, includes last contact)
14. `pdays` (Integer): number of days that passed by after the client was last contacted from a previous campaign (numeric; -1 means client was not previously contacted)
15. `previous` (Integer): number of contacts performed before this campaign and for this client
16. `poutcome` (Categorical): outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')

#### Social and economic context attributes
17. `emp.var.rate` (Integer): employment variation rate - quarterly indicator
18. `cons.price.idx` (Integer): consumer price index - monthly indicator
19. `cons.conf.idx` (Integer): consumer confidence index - monthly indicator
20. `euribor3m` (Integer): euribor 3 month rate - daily indicator
21. `nr.employed` (Integer): number of employees - quarterly indicator

#### Output variable (desired target)
22. `subscribed` (Binary): has the client subscribed a term deposit?

Source: [UCI Machine Learning Repository](https://archive.ics.uci.edu/dataset/222/bank+marketing)

```{r}
vis_dat(train)
skim(train)
```

```{r}
plot_ly(train, x = subscribed, type = 'histogram')

corrplot(cor(train[, c("X", "age", "duration", "campaign", "pdays", "previous", "emp.var.rate", "cons.price.idx", "cons.conf.idx", "euribor3m", "nr.employed")]), method="pie")
```

```{r}
plot_ly(train, x = job, y = age, type = 'box', color = job)

plot_ly(train, x = education, y = age, type = 'box', color = education)
```


```{r}
ord_edu <- train %>% count(education) %>%arrange(n)%>% pull(education)


eduResp <- ggplot(train, aes(x = factor(education, levels = ord_edu), fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  ylab("Proportion") +
  scale_fill_discrete(name = "Subscribed") +
  xlab("Education") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

eduFreq <- ggplot(as.data.frame(table(education)/sum(table(education))*100), aes(x = reorder(education, Freq), y = Freq)) +
  geom_bar(stat = "identity", color = "gray",  fill = "steelblue", alpha=0.9) +  
  coord_flip() +
  labs(title = "Education", x = "Education Level", y = "Count") +
  theme_minimal()

eduFreq / eduResp

ordine_poutcome <- train %>% count(poutcome) %>% arrange(n) %>%
  pull(poutcome)

poutcomeFreq <- ggplot(as.data.frame(table(train$poutcome) / length(train$poutcome) * 100),
                       aes(x = reorder(Var1, Freq), y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Distribution of Poutcome",
    x = "Outcome previous campaign",
    y = "Percentage (%)"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(legend.position = "none")

poutcomeResp <- ggplot(train, aes(x = factor(poutcome, levels = ordine_poutcome), fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  ylab("Proportion") +
  scale_fill_discrete(name = "Poutcome") +
  xlab("Outcome previous campaign") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal()

(poutcomeFreq / poutcomeResp) +
  plot_layout(axis_titles = 'collect')




ggplot(train, aes(age)) + geom_histogram(binwidth=4,position="fill",aes(fill=factor(subscribed)))+scale_fill_discrete(name="Subscribed")+ylab("proportion")+geom_hline(yintercept=0.5)

ordine_job <- train %>% count(job) %>%arrange(n)%>% pull(job)


jobFreq <- ggplot(as.data.frame(table(train$job) / length(train$job) * 100),
                   aes(x = reorder(Var1, Freq), y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", color = "gray",  fill = "steelblue", alpha=0.9) +
  coord_flip() +
  labs(
    title = "Distribution of job",
    x = "Outcome of the previous campaign",
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

jobResp <- ggplot(train, aes(x = factor(job, levels = ordine_job), fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(
    title = "Proportion by subscribed",
    x = "Default",
    y = "Proportion"
  ) +
  scale_fill_discrete(name = "Subscribed") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal()

jobFreq / jobResp


ggplot(train, aes(cons.price.idx)) + geom_histogram(binwidth=2,position="fill",aes(fill=factor(subscribed)))+scale_fill_discrete(name="Subscribed")+ylab("proportion")+geom_hline(yintercept=0.5)

ggplot(train, aes(cons.conf.idx)) + geom_histogram(binwidth=3,position="fill",aes(fill=factor(subscribed)))+scale_fill_discrete(name="Subscribed")+ylab("proportion")+geom_hline(yintercept=0.5)

ggplot(train, aes(euribor3m)) + geom_histogram(binwidth=3,position="fill",aes(fill=factor(subscribed)))+scale_fill_discrete(name="Subscribed")+ylab("proportion")+geom_hline(yintercept=0.5)


train$day_of_week <- factor(train$day_of_week,
                            levels = c("mon", "tue", "wed", "thu", "fri"),
                            ordered = TRUE)

dayFreq <- ggplot(as.data.frame(table(train$day_of_week)/length(train$day_of_week)*100), aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Distribution of Day of Week",
       x = "Last Contact Day of Week",
       y = "Percentage (%)")+
  theme_minimal()

dayResp <- ggplot(train, aes(x = day_of_week, fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  ylab("Proportion") +
  scale_fill_discrete(name = "Subscribed") +
  xlab("Last Contact Day of Week") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal() 

(dayFreq / dayResp) +
  plot_layout(axis_titles = 'collect')


ordine_month<-factor(train$month, 
                     levels = c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"), 
                     ordered = TRUE)

monthFreq <- ggplot(as.data.frame(table(ordine_month)/length(ordine_month)*100), aes(x = ordine_month, y = Freq)) +
  geom_bar(stat = "identity",color='gray', fill = "steelblue") +
  coord_flip() +
  labs(title = "Distribution of month",
       x = "Last contact month of year",
       y = "Percentage (%)")+
  theme_minimal()

monthResp <- ggplot(train, aes(x = ordine_month, fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  ylab("Proportion") +
  scale_fill_discrete(name = "month") +
  xlab("Last contact month of year") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal()

(monthFreq / monthResp) +
  plot_layout(axis_titles = 'collect')


ordine_previous <- train %>% count(previous) %>% arrange(n) %>%
  pull(previous) 

prevFreq <- ggplot(as.data.frame(table(train$previous)/length(train$previous)*100), aes(x = reorder(Var1,Freq), y = Freq)) +
  geom_bar(stat = "identity",color='gray', fill = "steelblue") +
  coord_flip() +
  labs(title = "Distribution of Previous",
       x = "Number of calls previous campain",
       y = "Percentage (%)")+
  theme_minimal()

prevResp <- ggplot(train, aes(x = factor(previous, levels = ordine_previous), fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  ylab("Proportion") +
  scale_fill_discrete(name = "Previous") +
  xlab("Number of calls previous campain") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal()

(prevFreq / prevResp) +
  plot_layout(axis_titles = 'collect')

train$emp_cat <- ifelse(train$emp.var.rate < 0, "Negative", "Positive or Zero")


ordine_emp <- train %>% count(emp_cat) %>% arrange(n) %>%
  pull(emp_cat)


empFreq <- ggplot(as.data.frame(table(ordine_emp)/length(ordine_emp)*100), aes(x = ordine_emp, y = Freq)) +
  geom_bar(stat = "identity",color='gray', fill = "steelblue") +
  coord_flip() +
  labs(title = "Distribution of Employment Variation (±)",
       x = "Employment Variation (±)",
       y = "Percentage (%)")+
  theme_minimal()

empResp <- ggplot(train, aes(x = factor(emp_cat, levels = ordine_emp), fill = factor(subscribed))) +
  geom_bar(position = "fill") +
  coord_flip() +
  ylab("Proportion") +
  scale_fill_discrete(name = "emp_cat") +
  xlab("Employment Variation (±)") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  theme_minimal()

(empFreq / empResp) +
  plot_layout(axis_titles = 'collect')

```

```{r}
ggpairs(train[, c("age", "duration", "campaign", "pdays", "previous", "emp.var.rate", "cons.price.idx", "cons.conf.idx", "euribor3m", "nr.employed")], columns = 1:10, 
                 lower = list(continuous = wrap("points", alpha = 0.5, color = "darkred", size=0.5)),
                 title='Scatterplot', axisLabels='none')
```
