## Exploratory Data Analysis

```{r, include=FALSE}
data <- read.csv("../data/raw/bank_marketing_train.csv")
```

```{r}
str(data)
```

1. `age` (Integer): age of the customer
2. `job` (Categorical): occupation
3. `marital` (Categorical): marital status
4. `education` (Categorical): education level
5. `default` (Binary): has credit in default?
6. `balance` (Integer): average yearly balance
7. `housing` (Binary): has housing loan?
8. `loan` (Binary): has personal loan?
9. `contact` (Categorical): contact communication type
10. `day_of_week` (Integer): last contact day of the week
11. `month` (Categorical): last contact month of year
12. `duration` (Integer): last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model
13. `campaign` (Integer): number of contacts performed during this campaign and for this client (numeric, includes last contact)
14. `pdays` (Integer): number of days that passed by after the client was last contacted from a previous campaign (numeric; -1 means client was not previously contacted)
15. `previous` (Integer): number of contacts performed before this campaign and for this client
16. `poutcome` (Categorical): outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
17. `subscribed` (Binary): has the client subscribed a term deposit?

Source: [UCI Machine Learning Repository](https://archive.ics.uci.edu/dataset/222/bank+marketing)

```{r}
vis_dat(data)
```
